---
title: "Data storage schema"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data storage

Remotely sensed multi-spectral data is represented by a combination of spatial position (pixel number and line number) and band. The _data storage scheme_ for multi-band images is classified into the following three types, described below:

1. Band sequential (BSQ): entire bands are stored on disk one after the other.
2. Band interleaved by line (BIL): bands are stored line-wise.
3. Band interleaved by pixel (BIP): for each band, gray values are stored in a pixel-wise manner.

So let's download a Sentinel 2 multi-spectral image and go over these schema's. We will use `sen2r` package to download the data. First install/load the package

```{r s1}
rm(list=ls(all=TRUE))
if(!require(sen2r))install.packages('sen2r')
library(sen2r)
```

Let us use a GUI to download sentinel 2 data.

```{r s2}
#sen2r()
```

Next lets save our login credentials by running the command `write_scihub_login("my_user", "my_password")` where 'my_user' is your username and 'my_password' the corresponding [scihub](https://scihub.copernicus.eu/) account password. If so wish you can use an interactive GUI using the command `sen2r()` but ensure to install the following packages `mapedit`, `shinyFiles`, `shinydashboard`, and `shinyWidgets`. 

```{r s3}
#write_scihub_login("my_user", "my_password")
# 
# path <- getwd()
# aoi <- paste0(path,"/data/JKUAT.json")
# out_dir <- paste0(path,"/data/")
# safe_dir <- dir.create(paste0(path,"/safe/"),recursive = TRUE)
# 
# out1 <- sen2r(
#   gui = FALSE,
#   extent = system.file(aoi, package = "sen2r"),
#   extent_name = 'JKUAT',
#   timewindow = c("2021-01-01", "2021-01-12"),
#   list_prods = "BOA",
#   list_rgb = "RGB832B",
#   path_l1c = safe_dir,
#   path_l2a = safe_dir,
#   path_out = out_dir,
#   clip_on_extent = TRUE,
#   extent_as_mask = TRUE,
#   overwrite =TRUE
# )
# 
# 
# safe_folder <- tempfile(pattern = "safe_")
# out_folder_1 <- tempfile(pattern = "Example1_")
# sel_tile_1 <- s2tiles_kaz[sample(nrow(s2tiles_kaz),1),] %>%
#   sf::st_transform(3857)
# sel_tile_1$tile_id
# ## [1] "40TET"
# sel_extent_1 <- sf::st_centroid(sel_tile_1) %>% st_buffer(5e3)
# out1 <- sen2r(
#   gui = FALSE,
#   extent = sel_extent_1,
#   timewindow = c("2018-02-21", "2018-03-02"),
#   list_rgb = "RGB432T",
#   path_l1c = safe_dir,
#   path_l2a = safe_dir,
#   path_out = out_dir,
#   log = log_path_1 <- tempfile()
# )

```



## References

Luigi Ranghetti and Mirco Boschetti and Francesco Nutini and Lorenzo Busetto (2020). sen2r: An R toolbox for automatically downloading and preprocessing Sentinel-2 satellite data. _Computers & Geosciences_, 139, 104473. doi: 10.1016/j.cageo.2020.104473, URL: https://sen2r.ranghetti.info.