---
title: "Data storage schema"
author: "Benson Kenduiywo"
date: "02/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data storage

Remotely sensed multi-spectral data is represented by a combination of spatial position (pixel number and line number) and band. The _data storage scheme_ for multi-band images is classified into the following three types, described below:

1. Band sequential (BSQ): entire bands are stored on disk one after the other.
2. Band interleaved by line (BIL): bands are stored line-wise.
3. Band interleaved by pixel (BIP): for each band, gray values are stored in a pixel-wise manner.

So let's download a Sentinel 2 multi-spectral image and go over these schema's. We will use `sen2r` package to download the data. First install/load the package

```{r s1}
rm(list=ls(all=TRUE))
if(!require(sen2r))install.packages('sen2r')
library(sen2r)
```

Let us use a GUI to donwload sentinel 2 data.

```{r s2}
sen2r()
```

Next lets save our login credentials by running the command `write_scihub_login("my_user", "my_password")` where 'my_user' is your username and 'my_password' the corresponding [scihub](https://scihub.copernicus.eu/) account password. If so wish you can use an interactive GUI using the command `sen2r()` but ensure to install the following packages `mapedit`, `shinyFiles`, `shinydashboard`, and `shinyWidgets`. 

```{r s2}
#write_scihub_login("my_user", "my_password")

path <- getwd()
aoi <- paste0(path,"/data/JKUAT.json")
out_dir <- paste0(path,"/data/")
safe_dir <- dir.create(paste0(path,"/safe/"),recursive = TRUE)

out1 <- sen2r(
  gui = FALSE,
  extent = system.file(aoi, package = "sen2r"),
  extent_name = 'JKUAT',
  timewindow = c("2021-01-01", "2021-01-12"),
  list_prods = "BOA",
  list_rgb = "RGB832B",
  path_l1c = safe_dir,
  path_l2a = safe_dir,
  path_out = out_dir,
  clip_on_extent = TRUE,
  extent_as_mask = TRUE,
  overwrite =TRUE
)


safe_folder <- tempfile(pattern = "safe_")
out_folder_1 <- tempfile(pattern = "Example1_")
sel_tile_1 <- s2tiles_kaz[sample(nrow(s2tiles_kaz),1),] %>%
  sf::st_transform(3857)
sel_tile_1$tile_id
## [1] "40TET"
sel_extent_1 <- sf::st_centroid(sel_tile_1) %>% st_buffer(5e3)
out1 <- sen2r(
  gui = FALSE,
  extent = sel_extent_1,
  timewindow = c("2018-02-21", "2018-03-02"),
  list_rgb = "RGB432T",
  path_l1c = safe_dir,
  path_l2a = safe_dir,
  path_out = out_dir,
  log = log_path_1 <- tempfile()
)

```


```{r}
library(RGISTools)
sres <- senSearch(startDate = as.Date("2021-01-01","%Y-%m-%d"),
endDate = as.Date("2021-01-10","%Y-%m-%d"),
platform = "Sentinel-2",
extent = sf::st_read("./data/JKUAT.kmz"),
product = "S2MSI1C",
username = "bensonkemboi",
password = "Roseflower#3700")

# filtering the path R094 where Navarre is located
names(sres)
sres.sen.R094 <- sres[grepl("R094", names(sres))]
names(sres.sen.R094)
# list the dates in sres
senGetDates(names(sres.sen.R094),format="%Y%j")
wdir <- file.path(tempdir(),"Path_for_downloading_folder")
# donwload the imagery
senDownload(searchres = sres.sen.R094,
username = "username",
password = "password",
AppRoot = wdir,
unzip = TRUE)
wdir.sen.unzip <- file.path(wdir,"Sentinel","unzip")
files.sen.unzip<-list.files(wdir.sen.unzip,
pattern = "\\TCI.jp2$",
full.names = TRUE,
recursive = TRUE)
img.sen.rgb<-stack(files.sen.unzip[1])
plotRGB(img.sen.rgb)
```


## References

Luigi Ranghetti and Mirco Boschetti and Francesco Nutini and Lorenzo Busetto (2020). sen2r: An R toolbox for automatically downloading and preprocessing Sentinel-2 satellite data. _Computers & Geosciences_, 139, 104473. doi: 10.1016/j.cageo.2020.104473, URL: https://sen2r.ranghetti.info.